---
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r echo = F}
knitr::opts_chunk$set(echo = T, warning = F)
options(dplyr.width = Inf)
```

# Estimates of flow-weighted mean concentrations

Load libraries and import data:
```{r, message = F, warning = F}
library(tidyverse)
library(readxl)
library(lubridate)

# metal data
dat <- read.csv('data/raw/Trace_Metals_Apr2019.csv', stringsAsFactors = F)

# watershed area data
wshed <- read_excel('data/raw/SDRS Revised StationID Table.xlsx', skip = 2)
```

Format watershed data:

* select only watershed name and area
* rename columns
* remove numbers from site, correct USJ to USJC to match metal sites
```{r}
wshed <- wshed %>% 
  select(`Station Code`, Area) %>% 
  na.omit %>% 
  rename(
    SITE = `Station Code`, 
    AREAKM2 = Area
  ) %>% 
  mutate(
    SITE = gsub('[0-9]*$', '', SITE),
    SITE = ifelse(SITE == 'USJ', 'USJC', SITE),
    AREAKM2 = as.numeric(AREAKM2)
  )
```

Format metal data:

* remove unnecessary columns
* separete site and storm numbers into two columns
* format date time by combining date and time
* join wshed area to data
* put in long format by consituent
* remove "DF" (dissolved fraction) variables
* concentration value as numeric
* add units column

```{r}
dat <- dat %>%
  select(-SIZE, -GEO) %>% 
  unite('DATETIME', DATE, TIME) %>% 
  mutate(
    STRM = gsub('^[a-z,A-Z]*', '', SITE),
    STRM = ifelse(STRM == '', 'x', STRM),
    SITE = gsub('[0-9]*$', '', SITE),
    DATETIME = mdy_hm(DATETIME, tz = 'Etc/GMT+8'),
    YR = year(DATETIME)
  ) %>% 
  left_join(wshed, by = 'SITE') %>% 
  gather('VAR', 'VAL', -SITE, -STRM, -DATETIME, -YR, -TYPE, -SEASON, -FLOWCMS, -AREAKM2) %>% 
  filter(!grepl('^DF', VAR)) %>% 
  mutate(
    VAL = as.numeric(VAL),
    UNIT = case_when(
      VAR %in% c('Cl', 'Hard', 'S04', 'TDS', 'TotAlk', 'TSS') ~ 'mg/L', 
      T ~ 'ug/L'
    )
  ) 
```

## All groupings

Group the metals data by aggregations to calculate:

```{r}
datgrp <- dat %>% 
  group_by(SITE, STRM, UNIT, AREAKM2, SEASON, YR, TYPE, VAR) %>% 
  nest
datgrp
```

A function that estimates flow-weighted mean concentrations:
```{r}
emc_fun <- function(UNIT, data){

  # get supporting data for calcs
  data <- data %>% 
    mutate(
      litersmin = FLOWCMS * 60 * 1000, 
      elapsed_min = c(1, difftime(DATETIME[-1], DATETIME[-nrow(.)], units = 'mins')),
      liters = litersmin * elapsed_min,
      litersresult = liters * VAL # either mg or ug depending on constituent
    )

  # mass emissions, conver to g
  massemis <- data$litersresult %>% na.omit %>% cumsum %>% max
  massemis_g <- case_when(
    UNIT %in% 'mg/L' ~ massemis * 1e-3, 
    UNIT %in% 'ug/L' ~ massemis * 1e-6
  )
  
  # elapsed days, emc
  dys <- difftime(max(data$DATETIME), min(data$DATETIME), units = 'days') %>% as.numeric
  liter_total <- sum(data$liters, na.rm = T)
  emc_gl <- massemis_g / liter_total

  out <- data.frame(massemis_g = massemis_g, liter_total, dys = dys, emc_gl = emc_gl)
  
  return(out)
  
}
```

Apply the function to each value in `data`:
```{r}
datgrp <- datgrp %>% 
  mutate(
    emc = pmap(list(UNIT, data), emc_fun)
  )
datgrp
```

Expand dataset into results;
```{r}
datgrp <- datgrp %>% 
  select(-data) %>% 
  unnest
datgrp
```

Estimate mass loadings as kg/day:
```{r}
datgrp <- datgrp %>% 
  mutate(
    massload_kgdy = massemis_g * 1e-3 / dys
  )
datgrp
```

Estimate flux rates as ratio between kg/yr and watershed area (km2):
```{r}
datgrp <- datgrp %>% 
  mutate(
    fluxrate_kgyrkm2 = massload_kgdy * 365 / AREAKM2
  )
datgrp
```

## Groupings by wet/dry, season only

Take `datgrp` object, remove site, storm grouping variable, sum mass emission and days by groupings for season, type, and trace metal/conventionals.
```{r}
datgrp_nosite <- datgrp %>% 
  dplyr::select(-SITE, -UNIT, -STRM, -emc_gl, -massload_kgdy, -fluxrate_kgyrkm2) %>% 
  group_by(AREAKM2, SEASON, TYPE, VAR) %>% 
  summarise(
    massemis_g = sum(massemis_g, na.rm = T),
    liter_total = sum(liter_total, na.rm = T), 
    dys = sum(dys, na.rm = T)
    )
datgrp_nosite
```

Calculate flow-weighted mean concentrations, estimate mass loadings as kg/day, and flux rates:
```{r}
datgrp_nosite <- datgrp_nosite %>% 
  mutate(
    emc_gl = massemis_g / liter_total,  
    massload_kgdy = massemis_g * 1e-3 / dys,
    fluxrate_kgyrkm2 =  massload_kgdy * 365 / AREAKM2
  )
datgrp_nosite
```

## Groupings by wet/dry only

Repeat the above analysis but remove season grouping.

```{r}
datgrp_nosite_noseas <- datgrp %>% 
  dplyr::select(-SEASON, -SITE, -UNIT, -STRM, -emc_gl, -massload_kgdy, -fluxrate_kgyrkm2) %>% 
  group_by(AREAKM2, TYPE, VAR) %>% 
  summarise(
    massemis_g = sum(massemis_g, na.rm = T),
    liter_total = sum(liter_total, na.rm = T), 
    dys = sum(dys, na.rm = T)
    )
datgrp_nosite_noseas
```

Calculate flow-weighted mean concentrations, estimate mass loadings as kg/day, and flux rates:
```{r}
datgrp_nosite_noseas <- datgrp_nosite_noseas %>% 
  mutate(
    emc_gl = massemis_g / liter_total,  
    massload_kgdy = massemis_g * 1e-3 / dys,
    fluxrate_kgyrkm2 =  massload_kgdy * 365 / AREAKM2
  )
datgrp_nosite_noseas
```
